Preventing Leakage of Distribution in Order Preserving Encryption Schemes \\ \vspace*{.3in} \small Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang \\ \small 6.857 Final Paper \\ \small May 15, 2013

### 2013-05-15 paper.tex
Finished product.

%--------Meta Data: Fill in your info------
\title{
  Preventing Leakage of Distribution in Order Preserving Encryption Schemes \\ \vspace*{.3in}
\small Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang \\
\small 6.857 Final Paper}
\date{\today}

@@

@@

The mOPE scheme by Popa et al (2013) is the first scheme to achieve ideal IND-OCPA security, and our schemes are built upon mOPE. Prior work on range query schemes all leak information about the underlying plaintext. Shi et al (2007) designed an encryption scheme that leaks the decrypted data for values that fall within the range query, but preserves the privacy of all other values. We chose to build on schemes on existing OPE schemes, instead of range query schemes, to guarantee that security of underlying plaintext even within the range queries.

Further, Popa et al (2013) provides a new security definition, same-time OPE security, which dictates that an adversary can learn the order only for values stored in the application at the same time, and is stronger than IND-OCPA. Our security guarantee of IND-CPDA, explained later in the paper, asks furthermore that the ciphertexts not leak the distribution of values.

@@

  RTOPE can be proven to be IND-CPDA, just like the DCS scheme.
  \begin{theorem}
    RTOPE is IND-CPDA secure.
  \end{theorem}
  \begin{proof}
    Like in the security proof for DCS, we shall show that encrypting $v = \{V_1, V_2, \ldots, V_{\kappa} \}$ and $w = \{W_1, W_2, \ldots, W_{\kappa}\}$ is information-theoretically the same, where $V_i, W_i$ are sequences of size $\lambda$. Again, we shall use induction on the number of repeated elements in either $V_i$ or $W_i$. \\ \\
When there are no repeated elements, we can use the same argument we used in the DCS proof. We note that the sequence $\{V_1, V_2, \ldots, V_{\kappa} \}$ has a bijection to a sequence of elements $v_{ij}$ of length $\kappa \lambda$, for $i \in \{1, 2, \ldots, \kappa\}$ and $j \in \{1, 2, \ldots, \lambda \}$. Winning this game then becomes equivalent to winning the IND-OCPA security game, since one can think of each $v_{ij}$ and $w_{ij}$ pair as a set of values that the adversary provides the client in each round $i$. Popa et al (2013) shows that their scheme, which is equivalent to the Distribution Confidentiality Scheme when there are no repetitions, is IND-OCPA secure. Thus, the base case of no repetitions is secure. \\ \\
Now, suppose that RTOPE is IND-CPDA secure when there are at most $k$ repetitions across $V_i$ and $W_i$. Now consider a sequence where there are $k+1$ repetitions across $V_i$ and $W_i$ for all $i$. Select two ciphertexts $c_1$ and $c_2$ of some repeated value $x$ and let us consider how it this pair of repetitions is inserted into the tree. Since RTOPE uses an encryption scheme which is IND-CCA for its ciphertexts, the adversary will not be able to tell whether $Dec(SK, c_1) = Dec(SK, c_2)$ with more than $1/2$ probability (by the definition of IND-CCA security). This means that the adversary cannot distinguish $c_1$ and $c_2$ as being repetitions with any advantage. Thus, a sequence of $k+1$ repetitions is information theoretically the same for the adversary as a sequence of $k$ repetitions, which we already know is IND-CPDA secure. Using induction, we have completed the proof.  
  \end{proof}

\section{Relationship Between DCS and RTOPE}
  In this section, we shall propose a notion of equivalence between two OPE schemes.

@@

In other words, for two schemes to be structurally expectation equivalent, there must exist a mapping $f$ from all of the possible OPE encodings from scheme $x$ to scheme $y$, where the probability of obtaining an encoding in $x$ is the same as the probability of obtaining the corresponding encoding in $y$ (based on the mapping). This must be the case for all possible encodings. \\ \\
This notion of structural expectation equivalence encompasses a number of useful features. In particular, notice that if $x$ is an OPE encoding scheme which is IND-CPDA secure, and $y$ is structurally expectation equivalent to $x$, then $y$ is also IND-CPDA secure. Due to this notion, it is sufficient to show the structural expectation equivalence of two schemes in order to link the IND-CPDA security of the schemes (the proof of this property is given in the theorem below). Another observation one can make is that the mapping in the structural expectation equivalence definition is bijective. Hence, if $x$ is structurally expectation equivalent to $y$, then $y$ is necessarily structurally expectation equivalent to $x$ because one can simply use $f^{-1}$ as the mapping, which necessarily exists because of $f$ being bijective.

\begin{theorem}
 If two OPE schemes $x,y$ are structurally expectation equivalent, then $x$ is IND-CPDA secure if and only if $y$ is IND-CPDA secure.
\end{theorem}
\begin{proof}
First, we observe that notion of equivalence is symmetric (since $f$ is a bijective mapping which is invertible), so that without loss of generality, we need only show that if $x$ is IND-CPDA secure, then $y$ is as well. So now we can suppose that $x$ is IND-CPDA secure and $x,y$ are equivalent schemes. \\ \\
Since $x$ is IND-CPDA secure, then for any two sequences $v = \{V_1, V_2, \ldots, V_{\kappa}\}$ and $w = \{W_1, W_2, \ldots, W_{\kappa} \}$ that the adversary provides, the adversary cannot correctly distinguish between $v$ and $w$ with non-neglible advantage. Now suppose by contradiction that $y$ is not IND-CPDA secure. Then there exist two sequences $v' = \{V_1', \ldots, V_{\kappa}'\}$ and $w' = \{W_1', \ldots, W_{\kappa}'\}$ of plaintexts that the adversary can provide to the client so that he can distinguish the two sequences with non-neglible advantage in the client-server interaction. This means that the adversary can distinguish the encoding $Encode_y(V_i') = \{Encode_y(V_{ij}' \}_{j=1}^{\lambda}$ for some $i$ from the encoding $Encode_y(W_i') = \{Encode_y(W_{ij}') \}_{j=1}^{\lambda}$. \\ \\
This means the adversary can also distinguish the mapping $Encode_x(V_i')$ from the encoding $Encode_x(W_i')$ by applying the bijective mapping $f$ (from the definition of equivalence) to obtain $Encode_x(V_i') = f(Encode_y(V_i'))$ and $Encode_x(W_i') = f(Encode_y(W_i'))$. This is a contradiction because it shows that $x$ is not IND-CPDA secure, which was one of our assumptions. Thus, we see by contradiction that $y$ must be IND-CPDA secure, which proves our result.
\end{proof}

@@

In this section, we shall investigate whether DCS and RTOPE are structurally expectation equivalent. To do this, we shall first examine some propositions.

@@

Since a subtree containing all the ciphertexts corresponding to a plaintext $x$ exists, we can now examine properties of this subtree. First, we show that the root of the smallest possible such subtree must be a ciphertext which gets decrypted to $x$.

@@

We have seen that the smallest subtree containing a repeated plaintext will be rooted at a ciphertext decrypting to that plaintext. This has a number of interesting applications. For instance, one can examine the series of decisions that are made in the RTOPE scheme by the client by examining this subtree and a smaller subtree in a recursive fashion. \\ \\
For DCS, one can see that moving left or right from the root element depends on the relative size of the root element with the new element being inserted or queried. For example, if $x$ is the value of a repeated plaintext, then DCS will mutate $p$ into $p + r$ where $r \in [0,1)$ are some random bits. Moving left at the root means that $p_1 < p_r$ where $p_r$ is the root value. This means that $p + r_1 < p + r_r$, which is equivalent to saying $r_1 < r_r$. Thus, moving left at the root is just the probability that the random number $r_1$ is smaller than the random number $r_r$ picked at the root. \\ \\
Now let us examine some series of moves down the tree when attempting to query or insert some node with plaintext value $x$, denoted by the sequence
  \begin{eqnarray}
    \{ m_0(x), m_1(x), \ldots, m_k(x) \}
  \end{eqnarray} \\ \\
We shall drop the $x$ for convenience and commonly refer to moves as $m_i$. We see that $m_i$ can take on two possibilities, 0 for left or 1 for moving right, and we can characterize the moves:
  \begin{eqnarray}
    m_i(x) = \left\{ \begin{array}{l l}
        0 & \textrm{if } p_i > x \\
        1 & \textrm{else}
    \end{array} \right.
  \end{eqnarray}\\ \\
Here $p_i$ is the underlying plaintext value of the node at the $i$th move.\\ \\
We shall proceed to show that the two schemes do not produce the same expected tree. \begin{theorem}
    DCS and RTOPE are not structurally expectation equivalent.
  \end{theorem}
  \begin{proof}
We shall prove this theorem by examining a sequence of moves when looking at the insertion of a repeated value $x$. In DCS, to move right so that $m_i = 1$, one must have $p_i < x + r$. To move left, one must have $p_i > x + r$. Although there exists a sequence of moves from the root to a leaf of the tree which includes every node in the tree, we need only examine the nodes which have a plaintext value equal to $x$ (before the addition of the random bit). This is because the traversal of DCS and RTOPE is the same for nodes which are not repeated (they share the normal binary tree traversal rule). Thus, we know that the moves in these cases will be equivalent and the probability of making the moves on non-repeated nodes will be equal. \\ \\
Thus, we can reduce our problem to considering a sequence of moves \{$m_i\}_{i=1}^k$ on the plaintext values which are equal to $x$. If we can show that the probability of making a particular sequence of moves on the repeated values is the same for DCS and RTOPE, then we will be done. Making any sequence of moves on repeated values in RTOPE will have probability $2^{-k}$ where $k$ is the number of repeated value in the sequence down the tree. This is becase one moves left or right with probability $\frac{1}{2}$. \\ \\
Now we shall examine the sequence of moves in DCS. Recall that we need only consider a sequence of moves $\{m_i\}_{i=1}^k$ which is performed on a single repeated plaintext value $x$. This means that in DCS, the plaintext values will be distinguished by the random bit $r_i$ so that $p_i = x + r_i$. Therefore, we need only compare $r_i$ to the value $r^*$, where $r^*$ is the random part of the new plaintext we are inserting. \\ \\
The probability that we make a series of moves $\{m_i\}_{i=1}^k$ is the probability that $r^*$ has the $j$ and $j+1$st smallest elements in the sequence of random values $\{ r_i\}_{i=1}^k$. Let $n = R(x)$ be the total number of repeated values of $x$. \\ \\
It is well known that the probability that the $j$th smallest element falling in the interval $[u, u + du]$ is given by:
    \begin{eqnarray}
      \frac{n!}{(j-1)!(n-j)!} u^{j-1} (1-u)^{n-j} du
    \end{eqnarray} \\ \\
Therefore, the probability that $r^*$ falls between the $j$th and $j+1$st value is just the difference between all potential values $u_1$ of the $j$th smallest element and all potential values $u_2$ of the $j+1$st element. This is:
    \begin{eqnarray}
      \int_{0}^1 \int_0^{u_2} Pr[U_2 = u_2] Pr[U_1 = u_1] Pr[u_1 < r^* < u_2] d_{u_1} d_{u_2}
    \end{eqnarray} \\ \\
Where we take $U_1$ as the $j$th smallest value and $U_2$ as the $j+1$st smallest value. Thus, the probability of $r^*$ falling in that interval is the integral over all possible values of $U_2$ and $U_1$, and the probabilitiy that it falls inside of the that interval. Since $Pr[u_1 < r^* < u_2] = (u_2 - u_1)$, we can write the above as:\\ \\
\begin{eqnarray}
      \int_{0}^1 \int_0^{u_2} \left( \frac{n!}{j!(n-j-1)!} u_2^j (1 - u_2)^{n-j-1} \right) \left( \frac{n!}{(j-1)!(n-j)!} u_1^{j-1} (1 - u_1)^{n-j} \right) (u_2 - u_1) d_{u_1} d_{u_2}
    \end{eqnarray} \\ \\
Letting $C(j) = \frac{n!}{j!(n-j-1)!} \frac{n!}{(j-1)!(n-j)!}$, and integrating the expression for $u_1$, we can obtain the following result:
    \begin{eqnarray}
      C(j) \int_0^1 u_2^j (1 - u_2)^{n-j-1} \left( \frac{(u_2 (n+1) - j) B(j, n-j+1) + u_2^j (1-u_2)^{n-j+1}}{n+1} \right) d u_2
    \end{eqnarray} \\ \\
The above is the probability that $r^*$ falls between the $j$th and $j+1$st values. Simulating this in the figure below, it is clear that this is not uniform for all values of $j$ (like RTOPE is). \\ \\
\begin{figure}[H]
      \centering
      \includegraphics[width=3in]{ball_result.png}
      \caption{Probability of $r^*$ falling between the $j$th and $j+1$st largest values in the sequence $\{r_i\}_{i=1}^n$}
    \end{figure}
  \end{proof}

\subsection{Convergence in Number of Re-encodings}

Although DCS and RTOPE are not structurally expectation equivalent, we can examine other properties between these two schemes. In particular, we can show that as $n \to \infty$, if $R_{DCS} / R_{RTOPE} = c$ for some constant $c$ where $R_{i}$ is the number of repetitions for scheme $i$, then the number of re-encodings in DCS and RTOPE will converge to the same number. 
\begin{theorem}
    As $n \to \infty$, if $R_{DCS} / R_{RTOPE} = c$ for some constant $c$, then the number of re-econdings in DCS and RTOPE will converge to the same number.
  \end{theorem}
  \begin{proof}
  \end{proof}

@@

To confirm our theoretical analysis, we implemented both the DCS and RTOPE encryption schemes on top of a simplified version of CryptDB. First, we created a framework that mimicked the necessary features of CryptDB, including the client-server interaction and full insertion and query functionality. We then made a testing framework which allowed us to easily change the number of inputs, the number of repeats in the inputs, and the choice of whether to use DCS or RTOPE. The results are summarized in the table below. \\ \\
\begin{table} % Add the following just after the closing bracket on this line to specify a position for the table on the page: [h], [t], [b] or [p] - these mean: here, top, bottom and on a separate page, respectively
\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c c} % The final bracket specifies the number of columns in the table along with left and right borders which are specified using vertical bars (|); each column can be left, right or center-justified using l, r or c. To specify a precise width, use p{width}, e.g. p{5cm}
\toprule % Top horizontal line
& \multicolumn{4}{c}{Number of OPE Re-Encodings} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-5} % Horizontal line spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
Type & Inputs & Repeats & Average & Std. Dev.\\ % Column names row
\midrule % In-table horizontal line
RTOPE & 100 & 0  & 7363 & 5003 \\
      & 100 & 10  & 4385 & 2051 \\
      & 100 & 20  & 3945 & 643\\
      & 100 & 40  & 2598 & 642\\
      \midrule
DCS   & 100 & 0  & 3776 & 791 \\
      & 100 & 10  & 4452 & 1059 \\
      & 100 & 20  & 3287 & 1191\\
      & 100 & 40  & 3754 & 582\\
      \midrule
RTOPE & 200 & 0  & 15078 & 6751 \\
      & 200 & 20  & 16586 & 13215\\
      & 200 & 40  & 15707 & 6924\\
      \midrule
DCS   & 200 & 0  & 11300 & 2512 \\
      & 200 & 20  & 12355 & 8704\\
      & 200 & 40  & 8274 & 1654\\
\bottomrule % Bottom horizontal line
\end{tabular}
\caption{Experimental results. N=10 for each set of inputs.} % Table caption, can be commented out if no caption is required
\label{tab:results_table} % A label for referencing this table elsewhere, references are used in text as \ref{label}
\end{table} \\ \\
Each iteration constructed an entire tree from random inputs fitting the input criteria. The averages and standard deviations describe the distribution of the number of OPE re-encodings required throughout the construction of the tree. We make several observations from the data above:  \\ \\
\begin{enumerate}
\item For large inputs, the average number of OPE re-encodings is roughly the same for both schemes. We found that even though the two schemes are not identical, they do have the same expectation for the shape of the resulting trees over a large number of iterations, so the averages will be the same.
\item The standard deviation in OPE re-encodings varies significantly between the two schemes. This has to do with the resulting shape of the trees formed for the same inputs - DCS is more likely to create an unbalanced tree, and thus the standard deviations tend to be higher than comparable RTOPE deviations.
\item
The average number of OPE re-encodings is roughly the same for any number of repeated items. DCS, each duplicate is converted into a unique number before the insertion phase. In RTOPE, each value is treated as if it were either larger or smaller than its duplicate during the insertion phase. In both cases, the server is not made aware of duplicate values, and thus their quantity has no bearing on the properties of the tree created by the server.
\end{enumerate}

@@

 We presented DCS and RTOPE, two order-preserving encryption schemes that achieve ideal IND-CPDA security, where an adversary learns nothing about the distribution of the plaintext values based on the ciphertexts, and only the order of the elements. DCS generates random bits to append to any plaintext value that is encrypted, which requires a modified querying scheme involving localized range queries. RTOPE handles repeated values differently, by appending repeated plaintexts as left or right children on the server's search tree with uniform probability. Again, we created a modified querying algorithm for RTOPE that requires recursive queries once one occurence of the plaintext query is found. \\ \\
We proved that DCS and RTOPE are IND-CPDA secure and introduced the notion of structurally expectation equivalence. We also proved the important theorem that if two OPE schemes are structurally expectation equivalent, then one is IND-CPDA secure if and only if the other is IND-CPDA secure as well. We then showed that though DCS and RTOPE do not satisfy this structurally expectation equivalence relationship, they are both IND-CPDA secure. \\ \\

>>>>

%--------Meta Data: Fill in your info------
\title{
  Preventing Leakage of Distribution in Order Preserving Encryption Schemes \\ \vspace*{.3in}
\small Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang \\
\small 6.857 Final Paper \\
\small May 15, 2013}
\date{}

@@

\abstract{Order Preserving Encryption (OPE) schemes have been heavily studied in the cryptography literature because of their potential application to database design. Unfortunately, few schemes have been able to achieve provable security and leak the minimum amount of information possible. Until Popa et al (2013), no OPE scheme achieved IND-OCPA security. However, even Popa et al's design potentially leaks the distribution of repeated ciphertexts. This paper provides two schemes which are both IND-OCPA secure and do not leak the distribution of ciphertexts. We introduce the notion of IND-CPDA security, and show that both of our schemes satisfy the notion. Finally, we provide experimental results on the effectiveness of these two schemes in practice.}

\newpage

\tableofcontents

\newpage

@@

The mOPE scheme by Popa et al (2013) is the first scheme to achieve ideal IND-OCPA security, and our schemes are built upon mOPE. Further, Popa et al (2013) provides a new security definition, same-time OPE security, which dictates that an adversary can learn the order only for values stored in the application at the same time, and is stronger than IND-OCPA. Our security guarantee of IND-CPDA, explained later in the paper, asks furthermore that the ciphertexts not leak the distribution of values. Thus, our two schemes provide both same-time OPE security and IND-CPDA security.

Prior work in the field of OPE schemes falls short of providing IND-OCPA, proposed by Boldyreva et al (2009). Boldyreva et al (2009) prove that no immutable scheme can achieve IND-OCPA, however, Popa et al (2013) prove that IND-OCPA is in fact achievable with mutable ciphertexts, as described in the procedure above. The security alternative that Boldyreva et al (2009) present is that of random order-preserving function (ROPF). ROPF asks that an OPE scheme look "as-random-as-possible", while preserving order, and Boldyreva et al (2009) provide an OPE scheme that satisfies the notion of ROPF based on a random order-preserving function. However, OPE schemes that fall under the ROPF notion have been shown to leak at least half of the plaintext bits, in addition to order.

Prior work on range query schemes all leak information about the underlying plaintext. Shi et al (2007) designed an encryption scheme that encrypts the summaries of encrypted data, such that an authority may decrypt summaries that fall within a specific range. This scheme leaks the decrypted data for values that fall within the range query, but preserves the privacy of all other values. There is a tradeoff here, as for range queries under OPE schemes, one peice of information can be leaked for every value - the order - while in Shi et al's (2007) scheme, decrypted values are leaked, but nothing beyond the range query. We chose to build our schemes upon mOPE, instead of range query schemes, to guarantee that security of underlying plaintext. However, it would be interesting to explore the layering of range query schemes with OPE schemes to protect both the order and plaintext values.

@@

  RTOPE can be proven to be IND-CPDA, just like the DCS scheme. \\

  \begin{theorem}
    RTOPE is IND-CPDA secure.
  \end{theorem}
  \begin{proof}
    Like in the security proof for DCS, we shall show that encrypting $v = \{V_1, V_2, \ldots, V_{\kappa} \}$ and $w = \{W_1, W_2, \ldots, W_{\kappa}\}$ is information-theoretically the same, where $V_i, W_i$ are sequences of size $\lambda$. Again, we shall use induction on the number of repeated elements in either $V_i$ or $W_i$.

When there are no repeated elements, we can use the same argument we used in the DCS proof. We note that the sequence $\{V_1, V_2, \ldots, V_{\kappa} \}$ has a bijection to a sequence of elements $v_{ij}$ of length $\kappa \lambda$, for $i \in \{1, 2, \ldots, \kappa\}$ and $j \in \{1, 2, \ldots, \lambda \}$. Winning this game then becomes equivalent to winning the IND-OCPA security game, since one can think of each $v_{ij}$ and $w_{ij}$ pair as a set of values that the adversary provides the client in each round $i$. Popa et al (2013) shows that their scheme, which is equivalent to the Distribution Confidentiality Scheme when there are no repetitions, is IND-OCPA secure. Thus, the base case of no repetitions is secure.

Now, suppose that RTOPE is IND-CPDA secure when there are at most $k$ repetitions across $V_i$ and $W_i$. Now consider a sequence where there are $k+1$ repetitions across $V_i$ and $W_i$ for all $i$. Select two ciphertexts $c_1$ and $c_2$ of some repeated value $x$ and let us consider how this pair of repetitions is inserted into the tree. Since RTOPE uses an encryption scheme which is IND-CCA for its ciphertexts, the adversary will not be able to tell whether $Dec(SK, c_1) = Dec(SK, c_2)$ with more than $1/2$ probability (by the definition of IND-CCA security). This means that the adversary cannot distinguish $c_1$ and $c_2$ as being repetitions with any advantage. Thus, a sequence of $k+1$ repetitions is information theoretically the same for the adversary as a sequence of $k$ repetitions, which we already know is IND-CPDA secure. Using induction, we have completed the proof.
  \end{proof}

\section{Relationship Between DCS and RTOPE}

  In this section, we shall propose a notion of equivalence between two OPE schemes and examine whether DCS and RTOPE are equivalent.

@@

In other words, for two schemes to be structurally expectation equivalent, there must exist a mapping $f$ from all of the possible OPE encodings from scheme $x$ to scheme $y$, where the probability of obtaining an encoding in $x$ is the same as the probability of obtaining the corresponding encoding in $y$ (based on the mapping). This must be the case for all possible encodings. \\ \\
This notion of structural expectation equivalence encompasses a number of useful features. In particular, notice that if $x$ is an OPE encoding scheme which is IND-CPDA secure, and $y$ is structurally expectation equivalent to $x$, then $y$ is also IND-CPDA secure. Due to this notion, it is sufficient to show the structural expectation equivalence of two schemes in order to link the IND-CPDA security of the schemes (the proof of this property is given in the theorem below). Another observation one can make is that the mapping in the structural expectation equivalence definition is bijective. Hence, if $x$ is structurally expectation equivalent to $y$, then $y$ is necessarily structurally expectation equivalent to $x$ because one can simply use $f^{-1}$ as the mapping, which necessarily exists because of $f$ being bijective. \\

\begin{theorem}
 If two OPE schemes $x,y$ are structurally expectation equivalent, then $x$ is IND-CPDA secure if and only if $y$ is IND-CPDA secure.
\end{theorem}
\begin{proof}
First, we observe that notion of equivalence is symmetric (since $f$ is a bijective mapping which is invertible), so that without loss of generality, we need only show that if $x$ is IND-CPDA secure, then $y$ is as well. So now we can suppose that $x$ is IND-CPDA secure and $x,y$ are equivalent schemes.

Since $x$ is IND-CPDA secure, then for any two sequences $v = \{V_1, V_2, \ldots, V_{\kappa}\}$ and $w = \{W_1, W_2, \ldots, W_{\kappa} \}$ that the adversary provides to scheme $x$, the adversary cannot correctly distinguish between $v$ and $w$ with non-neglible advantage. Now suppose by contradiction that $y$ is not IND-CPDA secure. Then there exist two sequences $v' = \{V_1', \ldots, V_{\kappa}'\}$ and $w' = \{W_1', \ldots, W_{\kappa}'\}$ of plaintexts that the adversary can provide to the client in scheme $y$ so that he can distinguish the two sequences with non-neglible advantage in the client-server interaction. This means that the adversary can distinguish the encoding $Encode_y(V_i') = \{Encode_y(V_{ij}') \}_{j=1}^{\lambda}$ for some $i$ from the encoding $Encode_y(W_i') = \{Encode_y(W_{ij}') \}_{j=1}^{\lambda}$.

This means the adversary can also distinguish the mapping $Encode_x(V_i')$ from the encoding $Encode_x(W_i')$ by applying the bijective mapping $f$ (from the definition of equivalence) to obtain $Encode_x(V_i') = f(Encode_y(V_i'))$ and $Encode_x(W_i') = f(Encode_y(W_i'))$. This is a contradiction because it shows that $x$ is not IND-CPDA secure, which was one of our assumptions. Thus, we see by contradiction that $y$ must be IND-CPDA secure, which proves our result.
\end{proof}

@@

In this section, we shall investigate whether DCS and RTOPE are structurally expectation equivalent. To do this, we shall first examine some propositions. \\

@@

Since a subtree containing all the ciphertexts corresponding to a plaintext $x$ exists, we can now examine properties of this subtree. First, we show that the root of the smallest possible such subtree must be a ciphertext which gets decrypted to $x$. \\

@@

We have seen that the smallest subtree containing a repeated plaintext will be rooted at a ciphertext decrypting to that plaintext. This has a number of interesting applications. For instance, one can examine the series of decisions that are made in the RTOPE scheme by the client by examining this subtree and a smaller subtree in a recursive fashion. \\ \\
For DCS, one can see that moving left or right from the root element depends on the relative size of the root element with the new element being inserted or queried. For example, if $x$ is the value of a repeated plaintext, then DCS will mutate $p$ into $p + r$ where $r \in [0,1)$ are some random bits. Moving left at the root means that $p_1 < p_r$ where $p_r$ is the root value. This means that $p + r_1 < p + r_r$, which is equivalent to saying $r_1 < r_r$. Thus, moving left at the root is just the probability that the random number $r_1$ is smaller than the random number $r_r$ picked at the root. \\ \\
Now let us examine some series of moves down the tree when attempting to query or insert some node with plaintext value $x$, denoted by the sequence
  \begin{eqnarray}
    \{ m_0(x), m_1(x), \ldots, m_k(x) \}
  \end{eqnarray}
We shall drop the $x$ for convenience and commonly refer to moves as $m_i$. We see that $m_i$ can take on two possibilities, 0 for left or 1 for moving right, and we can characterize the moves:
  \begin{eqnarray}
    m_i(x) = \left\{ \begin{array}{l l}
        0 & \textrm{if } p_i > x \\
        1 & \textrm{else}
    \end{array} \right.
  \end{eqnarray}

Here $p_i$ is the underlying plaintext value of the node at the $i$th move. We shall proceed to show that the two schemes do not produce the same expected tree. \\
\begin{theorem}
    DCS and RTOPE are not structurally expectation equivalent.
  \end{theorem}
  \begin{proof}
We shall prove this theorem by examining a sequence of moves when looking at the insertion of a repeated value $x$. In DCS, to move right so that $m_i = 1$, one must have $p_i < x + r$. To move left, one must have $p_i > x + r$. Although there exists a sequence of moves from the root to a leaf of the tree which includes every node in the tree, we need only examine the nodes which have a plaintext value equal to $x$ (before the addition of the random bit). This is because the traversal of DCS and RTOPE is the same for nodes which are not repeated (they share the normal binary tree traversal rule). Thus, we know that the moves in these cases will be equivalent and the probability of making the moves on non-repeated nodes will be equal.

Thus, we can reduce our problem to considering a sequence of moves \{$m_i\}_{i=1}^k$ on the plaintext values which are equal to $x$. By the propsition we proved above, this sequence of moves will be rooted at a ciphertext which is an encrypted value of $x$. Making any sequence of moves on repeated values in RTOPE will have probability $2^{-k}$ where $k$ is the number of repeated value in the sequence down the tree. This is becase one moves left or right with probability $\frac{1}{2}$.

Now we shall examine the sequence of moves in DCS. Recall that we need only consider a sequence of moves $\{m_i\}_{i=1}^k$ which is performed on a single repeated plaintext value $x$. This means that in DCS, the plaintext values will be distinguished by the random bit $r_i$ so that $p_i = x + r_i$. Therefore, we need only compare $r_i$ to the value $r^*$, where $r^*$ is the random part of the new plaintext we are inserting.

The probability that we make a series of moves $\{m_i\}_{i=1}^k$ is the probability that $r^*$ has the $j$ and $j+1$st smallest elements in the sequence of random values $\{ r_i\}_{i=1}^k$. Let $n = R(x)$ be the total number of repeated values of $x$.

It is well known that the probability that the $j$th smallest element falling in the interval $[u, u + du]$ is given by:
    \begin{eqnarray}
      \frac{n!}{(j-1)!(n-j)!} u^{j-1} (1-u)^{n-j} du
    \end{eqnarray}

Therefore, the probability that $r^*$ falls between the $j$th and $j+1$st value is just the difference between all potential values $u_1$ of the $j$th smallest element and all potential values $u_2$ of the $j+1$st element. This is:
    \begin{eqnarray}
      \int_{0}^1 \int_0^{u_2} Pr[U_2 = u_2] Pr[U_1 = u_1] Pr[u_1 < r^* < u_2] d_{u_1} d_{u_2}
    \end{eqnarray}

Where we take $U_1$ as the $j$th smallest value and $U_2$ as the $j+1$st smallest value. Thus, the probability of $r^*$ falling in that interval is the integral over all possible values of $U_2$ and $U_1$, and the probability that it falls inside of the that interval. Since $Pr[u_1 < r^* < u_2] = (u_2 - u_1)$, we can write the above as:
\begin{eqnarray}
      \int_{0}^1 \int_0^{u_2} \left( \frac{n!}{j!(n-j-1)!} u_2^j (1 - u_2)^{n-j-1} \right) \left( \frac{n!}{(j-1)!(n-j)!} u_1^{j-1} (1 - u_1)^{n-j} \right) (u_2 - u_1) d_{u_1} d_{u_2}
    \end{eqnarray}

Letting $C(j) = \frac{n!}{j!(n-j-1)!} \frac{n!}{(j-1)!(n-j)!}$, and integrating the expression for $u_1$, we can obtain the following result:
    \begin{eqnarray}
      C(j) \int_0^1 u_2^j (1 - u_2)^{n-j-1} \left( \frac{(u_2 (n+1) - j) B(j, n-j+1) + u_2^j (1-u_2)^{n-j+1}}{n+1} \right) d u_2
    \end{eqnarray}

The above is the probability that $r^*$ falls between the $j$th and $j+1$st values. Simulating this in the figure below, it is clear that this is not uniform for all values of $j$ (like RTOPE is).
    \begin{figure}[H]
      \centering
      \includegraphics[width=3in]{ball_result.png}
      \caption{Probability of $r^*$ falling between the $j$th and $j+1$st largest values in the sequence $\{r_i\}_{i=1}^n$}
    \end{figure}

  We have therefore shown that DCS and RTOPE are not structurally expectation equivalent.
  \end{proof}

@@

To confirm our theoretical analysis, we implemented both the DCS and RTOPE encryption schemes on top of a simplified client server interface (like the one Popa et al (2013) used). First, we created a framework that mimicked the necessary features of an mOPE, including the client-server interaction and full insertion and query functionality. We then made a testing framework which allowed us to easily change the number of inputs, the number of repeats in the inputs, and the choice of whether to use DCS or RTOPE. The results are summarized in the table \ref{tab:results_table}.

\begin{table} % Add the following just after the closing bracket on this line to specify a position for the table on the page: [h], [t], [b] or [p] - these mean: here, top, bottom and on a separate page, respectively
\centering % Centers the table on the page, comment out to left-justify
\begin{tabular}{l c c c c} % The final bracket specifies the number of columns in the table along with left and right borders which are specified using vertical bars (|); each column can be left, right or center-justified using l, r or c. To specify a precise width, use p{width}, e.g. p{5cm}
\toprule % Top horizontal line
& \multicolumn{4}{c}{Number of OPE Re-Encodings} \\ % Amalgamating several columns into one cell is done using the \multicolumn command as seen on this line
\cmidrule(l){2-5} % Horizontal line spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
Type & Inputs & Repeats & Average & Std. Dev.\\ % Column names row
\midrule % In-table horizontal line
RTOPE & 100 & 0  & 7363 & 5003 \\
      & 100 & 10  & 4385 & 2051 \\
      & 100 & 20  & 3945 & 643\\
      & 100 & 40  & 2598 & 642\\
      \midrule
DCS   & 100 & 0  & 3776 & 791 \\
      & 100 & 10  & 4452 & 1059 \\
      & 100 & 20  & 3287 & 1191\\
      & 100 & 40  & 3754 & 582\\
      \midrule
RTOPE & 200 & 0  & 15078 & 6751 \\
      & 200 & 20  & 16586 & 13215\\
      & 200 & 40  & 15707 & 6924\\
      \midrule
DCS   & 200 & 0  & 11300 & 2512 \\
      & 200 & 20  & 12355 & 8704\\
      & 200 & 40  & 8274 & 1654\\
\bottomrule % Bottom horizontal line
\end{tabular}
\caption{Experimental results. N=10 for each set of inputs.} % Table caption, can be commented out if no caption is required
\label{tab:results_table} % A label for referencing this table elsewhere, references are used in text as \ref{label}
\end{table}

Each iteration constructed an entire tree from random inputs fitting the input criteria. The averages and standard deviations describe the distribution of the number of OPE re-encodings required throughout the construction of the tree. We make several observations from the data:

\begin{enumerate}
\item For large inputs, the average number of OPE re-encodings is roughly the same for both schemes. We found that even though the two schemes are not identical, they do have the same expectation for the shape of the resulting trees over a large number of iterations, so the averages will be the same.
\item The standard deviation in OPE re-encodings varies significantly between the two schemes. This has to do with the resulting shape of the trees formed for the same inputs - DCS is more likely to create an unbalanced tree, and thus the standard deviations tend to be higher than comparable RTOPE deviations.
\item
The average number of OPE re-encodings is roughly the same for any number of repeated items. In DCS, each duplicate is converted into a unique number before the insertion phase. In RTOPE, each value is treated as if it were either larger or smaller than its duplicate during the insertion phase. In both cases, the server is not made aware of duplicate values, and thus their quantity has no bearing on the properties of the tree created by the server.
\end{enumerate}

@@

 We presented DCS and RTOPE, two order-preserving encryption schemes that achieve ideal IND-CPDA security, where an adversary learns nothing about the distribution of the plaintext values based on the ciphertexts, and only the order of the elements. DCS generates random bits to append to any plaintext value that is encrypted, which requires a modified querying scheme involving localized range queries. RTOPE handles repeated values differently, by appending repeated plaintexts as left or right children on the server's search tree with uniform probability. Again, we created a modified querying algorithm for RTOPE that requires recursive queries once one occurence of the plaintext query is found.

We proved that DCS and RTOPE are IND-CPDA secure and introduced the notion of structurally expectation equivalence. We also proved the important theorem that if two OPE schemes are structurally expectation equivalent, then one is IND-CPDA secure if and only if the other is IND-CPDA secure as well. We then showed that DCS and RTOPE do not satisfy this structurally expectation equivalence relationship.

We noted that though the schemes are not structurally expectation equivalent, the number of OPE re-encodings will converege to the same number, given identical inputs. This notion was supported by our experimental results, which we tested on a range of inputs.
